## Data Exploration of Baseball (Spring)

#### Matthew Neus

### Introduction

This notebook serves as an example of basic data analysis of Maryland high school sports records.This data focuses on baseball records in the spring.

### Loading Libraries


```{r}
library(tidyverse)
```

### Loading Data

```{r}
bballlogs <- read_csv("https://raw.githubusercontent.com/maryland-high-school-sports/record-books/refs/heads/main/spring/baseball/championships.csv")
```

### Basic Aggregation

Some examples using group_by() and summarize()

```{r}
champion_data <- bballlogs |> filter(Status == "Champion")


championship_counts <- champion_data |>
  group_by(Team) |>
  summarize(championships_won = n()) |>
  arrange(desc(championships_won))
```

```{r}
finalist_data <- bballlogs |> filter(Status == "Finalist")


finalist_counts <- finalist_data |>
  group_by(Team) |>
  summarize(finals_apps = n()) |>
  arrange(desc(finals_apps))
```

```{r}
coach_data <- bballlogs |>
  group_by(Coach) |>
  summarise(CoachApps = n()) |>
arrange(desc(CoachApps))
```

### Visualizations

```{r}
top_champions <- coach_data |> top_n(10, CoachApps)

ggplot(top_champions, aes(x = reorder(Coach, -CoachApps), y = CoachApps)) +
  geom_bar(stat = "identity", fill = "grey") +
  labs(title = "Bill Dunham, Jim Steele and Pete Moylan Know How to Win",
       x = "Team", y = "Number of Championshipship Appearances") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
top_champions <- championship_counts |> top_n(10, championships_won)

ggplot(top_champions, aes(x = reorder(Team, -championships_won), y = championships_won)) +
  geom_bar(stat = "identity", fill = "grey") +
  labs(title = "Loch Raven, St. John's Swings for the Fences",
       x = "Team", y = "Number of Championships") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Notes/Caveats/Problems

I did not have any issues with the status of teams depending on if they were a champion or finalist when working with the data. I made a bar chart to show that Loch Raven and St. John's have the most championships, and that three coaches get to the finals more than most of their opposition.

The only problem I ran into was in the fact checking portion. Some schools were listed as champions in years they didn't win a championship, and the same thing happened for some teams listed as finalists.