## Data Exploration of Dual Meet Wrestling (Winter)

#### Pablo Suarez

### Introduction

This notebook serves as an example of basic data analysis of Maryland high school sports records, specifically for dual meet wrestling.

### Loading Libraries


```{r}
library(tidyverse)
library(ggplot2)
library(ggalt)
library(dplyr)
library(gt)
library(cowplot)
library(ggrepel)
library(knitr)
```

### Loading Data

```{r}

championships <- read_csv("https://raw.githubusercontent.com/maryland-high-school-sports/record-books/refs/heads/main/winter/dual-meet-wrestling/championships.csv")

championships
```

### Basic Aggregation

Some examples using group_by() and summarize()

```{r}

titles <- championships |>
  group_by(Team) |>
  filter(Status == "Champion") |>
  summarise(count = n()) |>
  arrange(desc(count))

```

```{r}

championships |>
  group_by(Team) |>
  filter(Status == "Finalist") |>
  summarise(count = n()) |>
  arrange(desc(count))

```

```{r}

coaches <- championships |>
  filter(Status == "Champion") |>
  group_by(Coach, Team) |>
  summarise(count = n(), .groups = "drop") |>
  arrange(desc(count))

```



### Visualizations

Use the aggregations (or other dataframes you create) to make some simple charts

```{r}

ggplot(titles, aes(x = factor(Team), y = count)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "Damascus Sets the Standard for Powerhouse Programs",
       subtitle = "MPSSAA Championship Dual Meet Wrestling Schools",
       x = "Winning Schools",
       y = "Number of Championships",
       caption = "Data: MPSSAA") +
  scale_y_continuous(breaks = seq(0, max_count, by = 1)) +  
  theme_minimal() +
  coord_flip()

```

```{r}
coaches |> 
  gt() |> 
  cols_label(
    count = "Championships Won"
  ) |>
  tab_header(
    title = "The MPSSAA Dual Meet Wrestling Royalty",
    subtitle = "Coaches with the most titles in MPSSAA history."
  )
```


### Notes/Caveats/Problems

Describe any issues you had with the data and any caveats a user should know.

Naming Issues: In the first block of basic aggregation code, I noticed that one of the teams that appears in the output is listed as "SouthernN/AGarrett." The next code block saw a team appear with the name "NorthernN/ACalvert." I also noted other schools named as "NortheastN/AAA" and "ChesapeakeN/AAA."

