## Data Exploration of Swimming and Diving (Winter)

<<<<<<< HEAD
#### Mauriah Allen
=======
#### Marwa Barakat
>>>>>>> fd9fe7c277eb5e69ec49e83dc01e81b71e2c385a

### Introduction

This notebook serves as an example of basic data analysis of Maryland high school sports records.

### Loading Libraries

<<<<<<< HEAD

```{r}
library(tidyverse)
=======
```{r}
library(tidyverse)
library(janitor)
library(ggplot2)
>>>>>>> fd9fe7c277eb5e69ec49e83dc01e81b71e2c385a
```

### Loading Data

```{r}
<<<<<<< HEAD
swimanddive <- read_csv("championships.csv")
=======
championships <- read_csv('championships.csv') |> clean_names()
>>>>>>> fd9fe7c277eb5e69ec49e83dc01e81b71e2c385a
```

### Basic Aggregation

Some examples using group_by() and summarize()

```{r}
<<<<<<< HEAD
championswimdata <- swimanddive |> group_by(school, Status = "Champion") |> summarise(count=n())
```

```{r}
champover4 <- championswimdata |> filter(count > 4) 
=======
championships |>
  group_by(school) |>
  summarise(
    total_finals = n()
  ) |> arrange(desc(total_finals))
```

```{r}
swim_by_gender <- championships |>
  filter(status=='Champion') |>
  group_by(school, gender) |>
  summarise(
    total_champs = n()
  ) |> arrange(desc(total_champs))

swim_by_gender
>>>>>>> fd9fe7c277eb5e69ec49e83dc01e81b71e2c385a
```

### Visualizations

```{r}
ggplot() + 
  geom_bar(
<<<<<<< HEAD
    data=champover4, 
    aes(x=reorder(school, count),
        weight=count))  + 
  labs(
    title="Top Swim and Dive Programs By Championship",
    subtitle = "Poolesville is exceding the rest of the schools championship record by double ",
    x="School", 
    y="Championships") + 
  theme_minimal() + 
  coord_flip()
```



### Notes/Caveats/Problems

I had no problems with this data set, the data is imported from the Maryland State record book. In the state of Maryland, there are both boy and girl swim championship data and individual data. 

=======
    data=swim_by_gender, 
    aes(x=reorder(school, total_champs), weight=total_champs, fill=gender)) + 
  coord_flip() +
  labs(
    title = 'Championship Titles per School',
    x = 'School',
    y = 'Total Championships'
  )
```

### Notes/Caveats/Problems

I had no issues analyzing this data with the methods I used.
>>>>>>> fd9fe7c277eb5e69ec49e83dc01e81b71e2c385a
