## Data Exploration of Swimming and Diving (Winter)

#### Marwa Barakat

### Introduction

This notebook serves as an example of basic data analysis of Maryland high school sports records.

### Loading Libraries

```{r}
library(tidyverse)
library(janitor)
library(ggplot2)
```

### Loading Data

```{r}
championships <- read_csv('championships.csv') |> clean_names()
```

### Basic Aggregation

Some examples using group_by() and summarize()

```{r}
championships |>
  group_by(school) |>
  summarise(
    total_finals = n()
  ) |> arrange(desc(total_finals))
```

```{r}
swim_by_gender <- championships |>
  filter(status=='Champion') |>
  group_by(school, gender) |>
  summarise(
    total_champs = n()
  ) |> arrange(desc(total_champs))

swim_by_gender
```

### Visualizations

```{r}
ggplot() + 
  geom_bar(
    data=swim_by_gender, 
    aes(x=reorder(school, total_champs), weight=total_champs, fill=gender)) + 
  coord_flip() +
  labs(
    title = 'Championship Titles per School',
    x = 'School',
    y = 'Total Championships'
  )
```

### Notes/Caveats/Problems

I had no issues analyzing this data with the methods I used.
