## Data Exploration of Volleyball (Fall)

<<<<<<< HEAD
#### Mauriah Allen
=======
#### Sam Benning
>>>>>>> fd9fe7c277eb5e69ec49e83dc01e81b71e2c385a

### Introduction

This notebook serves as an example of basic data analysis of Maryland high school sports records.

### Loading Libraries

<<<<<<< HEAD

```{r}
library(tidyverse)
=======
```{r}
library(tidyverse)
library(ggplot2)
>>>>>>> fd9fe7c277eb5e69ec49e83dc01e81b71e2c385a
```

### Loading Data

```{r}
<<<<<<< HEAD
volleyball <- read_csv("championships.csv")
=======
vb <- read.csv("https://raw.githubusercontent.com/maryland-high-school-sports/record-books/refs/heads/main/fall/volleyball/championships.csv")

head(vb) #displays a preview of the data
>>>>>>> fd9fe7c277eb5e69ec49e83dc01e81b71e2c385a
```

### Basic Aggregation

Some examples using group_by() and summarize()

<<<<<<< HEAD
```{r}
championdata <- volleyball |> group_by(School, Status = "Champion") |> summarise(count=n())
```
```{r}
champover10 <- championdata |> filter(count > 10) 
```

### Visualizations

```{r}
ggplot() + 
  geom_bar(
    data=champover10, 
    aes(x=reorder(School, count),
        weight=count))  + 
  labs(
    title="Top Volleyball Programs By Championship",
    subtitle = "With 25 championships, Williamsport is leading the state",
    x="School", 
    y="Championships") + 
  theme_minimal() + 
  coord_flip()
```


### Notes/Caveats/Problems

I had no problems with this data set, the data is imported from the Maryland State record book. In the state of Maryland, there are no boy volleyball championship data or individual data, so this data is just for the high school girls. 

=======
Finding which school has won the most championships:

```{r}
vb |> group_by(School, Status) |> filter(Status == "Champion") |>
  summarize(
    count = n()
  ) |>
  arrange(desc(count))
```

Williamsport has won 16 championships in volleyball since the beginning of these records.

A more complex analysis: Which team has won the most back-to-back championships?

```{r}
vb |> group_by(School, Status) |> filter(Status == "Champion") |>
mutate(consecutive = Year - lag(Year) == 1) %>% #check for consecutive years using lag
  summarize(
    back_to_back = sum(consecutive, na.rm = TRUE) #count consecutive occurrences
  ) |>
  arrange(desc(back_to_back))
```

Centennial has won in consecutive years 9 times. Let's see which years they did it in:

```{r}
vb |> select(School, Year, Status) |> filter(School == "Centennial" & Status == "Champion")
```

They have had a number of streaks with more than just back-to-back titles. Wow!

### Visualizations

Let's say we wanted to visualize Centennial's dominance in volleyball over the years. We can create a stacked bar chart that shows how many times they succeeded in the state tournament, and organize it by each decade.

```{r}
#Create a new variable `Era` to show Centennial's success through each decade:
centennial <- vb |> filter(School == "Centennial") |> mutate(Era = case_when(
    Year >= 1970 & Year < 1980 ~ "1970s",
    Year >= 1980 & Year < 1990 ~ "1980s",
    Year >= 1990 & Year < 2000 ~ "1990s",
    Year >= 2000 & Year < 2010 ~ "2000s",
    Year >= 2010 & Year < 2020 ~ "2010s",
    Year >= 2020 & Year < 2030 ~ "2020s",
    TRUE ~ "Other"  #For unexpected values
  )) |>
    mutate(
    Era = factor(Era, levels = c("1970s", "1980s", "1990s", "2000s", "2010s", "2020s")) #This edit allows the era's to be in order when we plot them
  ) |>
  group_by(Era, Status) %>% 
  summarise(Count = n()) #This is how we save the data to a dataframe that we will use for our chart

#Plot the stacked bar chart using ggplot
ggplot() + 
  geom_bar(
    data=centennial, 
    aes(x=Era, weight=Count, fill=Status)) + 
  coord_flip() +
  labs(title = "Centennial's Championship Dominance Spread Across Two Decades")
```

### Notes/Caveats/Problems

None at the moment!
>>>>>>> fd9fe7c277eb5e69ec49e83dc01e81b71e2c385a
