## Data Exploration of Field Hockey (Fall)

#### Marwa Barakat

### Introduction

This notebook serves as an example of basic data analysis of Maryland high school sports records.

### Loading Libraries

```{r}
library(tidyverse)
library(ggplot2)
library(janitor)
```

### Loading Data

```{r}

championships <- read_csv('championships.csv') |> clean_names()
```

### Basic Aggregation

Some examples using group_by() and summarize()

```{r}
championships |>
  group_by(team) |>
  summarise(
    total_finals = n()
  ) |> arrange(desc(total_finals))
```

```{r}
total_champs <- championships |>
  filter(status=='Champion') |>
  group_by(team) |>
  summarise(
    total_champ = n()
  ) |> arrange(desc(total_champ)) 

total_champs
```

### Visualizations

```{r}
ggplot() + 
  geom_bar(
    data=total_champs, 
    aes(x=reorder(team, total_champ), weight=total_champ, fill=total_champ)) + 
  coord_flip() +
  labs(
    title = 'Championship Titles per School',
    x = 'School',
    y = 'Total Championships'
  )
```

### Notes/Caveats/Problems

It's difficult to do any analysis using the Classes of teams because most rows don't have a class listed. Also, the schools are all labeled in all caps font, which looks busy on graphs.
